---
title: "Machine Learning for Beginners"
format: 
  gfm:
    df-print: kable
self-contained: true
---

Machine Learning (ML) models are used in various applications, affecting societies directly or indirectly in daily life. The Machine learning (ML) field connects computer science and statistics, allowing computers to use collected data or past experience to solve unseen problems. It’s a must-have skill for every analyst and data scientist. It is also an important skill for anyone who wants to draw conclusions and make scientific suggestions based on collected raw data. A wide range of successful ML applications exist, including systems to predict customer behavior, support decision-making, recognize faces or spoken speech, optimize robot behavior, and extract knowledge from biological data. During this Workshop, participants will learn how to write and train ML models that can help in their field of study. Attendees will have the chance to learn the basics of Python and how to use it to build machine-learning models. In this Workshop, we will use Google Colab as our IDE, but students are free to use their favorite IDE.

The materials on this website are [CC-BY-4.0](https://creativecommons.org/licenses/by/4.0/) licensed.

![cc](https://mirrors.creativecommons.org/presskit/icons/cc.svg) 
![by](https://mirrors.creativecommons.org/presskit/icons/by.svg)


## Lectures
```{r}
#| echo: false
readxl::read_xlsx("mlb_24.xlsx") |>
  dplyr::mutate(Topic = ifelse(!is.na(Slides), glue::glue("[{Title}]({Slides})"), Title), .keep = "unused") |>
  dplyr::relocate(Topic, .after = Date) |> 
  dplyr::filter(Type == "lecture") |>
  dplyr::select(Day, Date, Topic)
```

## Exercises
```{r}
#| echo: false
readxl::read_xlsx("mlb_24.xlsx") |>
  dplyr::mutate(Topic = ifelse(!is.na(Title), glue::glue("{Title}"), Title), .keep = "unused") |>
  dplyr::mutate(HTML = ifelse(!is.na(htmlURL), glue::glue("[{HTML}]({htmlURL})"), HTML), .keep = "unused") |>
  dplyr::mutate(Notebook = ifelse(!is.na(NotebookURL), glue::glue("[{Notebook}]({NotebookURL})"), Notebook), .keep = "unused") |> 
  dplyr::filter(Type == "exercise") |>
  dplyr::select(Day, Date, Topic, HTML, Notebook)
```

## Answers
```{r}
#| echo: false
readxl::read_xlsx("mlb_24.xlsx") |>
  dplyr::mutate(Topic = ifelse(!is.na(Title), glue::glue("{Title}"), Title), .keep = "unused") |>
  dplyr::mutate(HTML = ifelse(!is.na(htmlURL), glue::glue("[{HTML}]({htmlURL})"), HTML), .keep = "unused") |>
  dplyr::mutate(Notebook = ifelse(!is.na(NotebookURL), glue::glue("[{Notebook}]({NotebookURL})"), Notebook), .keep = "unused") |> 
  dplyr::filter(Type == "solution") |>
  dplyr::select(Day, Date, Topic, HTML, Notebook)
```

## Google Colab
During the workshop, we will use [Google Colab](https://colab.research.google.com/) as our IDE. We will start by creating a new notebook. 


#### Checking the Python version and installing different versions 

For checking the version of Python that is already installed, we run the command:

```
!python --version
```

If we need to install a different version of Python (e.g. 3.7), we can use:

```
!apt-get install python3.7
```

#### Installing more Libraries

Most of the important libraries that we may need are already installed in the Google Colab environment. However, if you would like to install additional libraries, you can use the `pip` command. For example, to install `py_stringmatching`, you can use the command:

```
!pip install py_stringmatching
```


**Note:** when running system commands in Google colab, we use `!` before the command. This is common for running system commands in any notebook environment (you may also use the % symbol `%pip install py_stringmatching`).



#### Running simple Python code

Run the following Python code and explain what the code is doing:

```
t, f = True, False
print(t and f) 
print(t or f)  
print(not t)   
```
---
## About the Speaker

Abdulhakim Qahtan is an assistant professor at the Data Intensive Systems (DIS) Group, 
Information and Computing Sciences Department. 
Before joining Utrecht University, he worked as a Postdoctoral Researcher 
at Qatar Computing Research Institute (QCRI), 
Hamad Bin Khalifa University (HBKU), Qatar (2016-2019). 
Dr. Qahtan earned his PhD degree from the Machine Intelligence & kNowledge Engineering (MINE) Lab 
at King Abdullah University of Science and Technology (KAUST) (2016).  
He completed his B.S. and M.S. in Computer Science at Cairo University, 
Egypt and King Fahd University of Petroleum and Minerals (KFUPM), Saudi Arabia, 
respectively. He worked as a teaching assistant at Taiz university, 
Yemen and a lecturer at KFUPM, Saudi Arabia.  
His current research focuses on data cleaning, data stream mining, 
time series analysis, algorithmic fairness and explainable machine learning.

Website: https://www.uu.nl/staff/AAAQahtan

LinkedIn: https://www.linkedin.com/in/hakimqahtan/

Email: a.a.a.qahtan(at)uu.nl




